%!TEX root=../slr.tex
\section{Introduction}
\label{sec:introduction}

% Contexto;
% Problema;
% Objectivos;
% Métodos;
% Resultados / contribuição;
% Conclusão;

This article comes as a response to the necessary State of the Art
search that was required for project ATMOS, which was a Portuguese EU
funded project that aimed to develop a miniaturised spectroscopy system
for atmospheric analysis and trace gas mapping using Differential
Optical Absorption Spectroscopy (DOAS), with tomographic capabilities.
FCT NOVA participated in this project as part of a consortium which also
included Compta, one of the oldest IT groups in Portugal. The project
aimed to develop a miniaturised and highly mobile spectroscopic system
with tomographic capabilities, designed to map a small geographic region
with respect to the concentration of a set of atmospheric trace gases,
like NO\textsubscript{2} or O\textsubscript{3}, using a technique called
Differential Optical Absorption Spectroscopy (DOAS).

DOAS is one of the most prominent methods for analysing and quantifying
atmospheric chemistry, namely in what concerns trace gas concentrations.
The technique, developed during the 70s by Perner and
Platt~\cite{Perner1976}, was popularised in the following decades by its
use in detecting Ozone, Nitrogen Oxydes and studies of cloud radiative
transport. DOAS is a type of absorption spectroscopy, which uses a
clever mathematical and physical observation to overcome the
difficulties of spectral measurement in the open atmosphere.

Through the setting of very careful geometric considerations, it is
possible to combine DOAS with tomographic reconstruction methods in
order to assemble a map of the gaseous concentrations in a given
geographic region. Tomography is the process of reconstructing an image
through projections obtained by subjecting a given target (in our case,
the atmosphere) to being traversed by any kind of penetrating or
reflecting wave, which in our case is visible light.

With this study, we have intended to capture the current literary
landscape surrounding the usage of tomographic DOAS, assessing this
technique's technological status. For this purpose, we have employed a
review methodology originary from Evidence Based Medicine. This method,
which has migrated to engineering through Software Engineering, is
called a Systematic Mapping Study (MS). It provides a framework that
allows researchers to produce detailed and systematic search protocols,
which are used to catalogue literature information and identify research
gaps within a determined subject.

The search procedure that we have defined was carefully engineered to
cover all tomographic DOAS research relevant to urban, rural or
industrial environments, with the main goal of finding a new
investigative path to follow. Through it, we were able to find several
different applications, all pertaining to scientific research, which are
similar regarding physical principle, but differ in objectives,
equipment assembly, algorithms, software, and geometry. Most of the
selected papers use an active DOAS principle for measuring atmospheric
chemical concentrations, and all of them are either fixed or of low
mobility. 

The rest of this paper has the following structure:
Section~\ref{sec:background} presents the context within which this
study was written; Section~\ref{sec:methods} describes how we have
planned to perform the study and the methods we have used in doing it;
Section~\ref{sec:conduction} describes the application of the said
methods in the pursuit of our goals and presents the results we have
obtained, as well as our evaluation of our processes;
Section~\ref{sec:conclusions} shows our conclusions and what we think
might be retained from reading this paper.
