%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% memstore.tex
%% NOVA thesis documentation file
%%
%% Detailed explanation of memstore.sty functionality.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE memstore.tex}

\chapter{Explanation of memstore.sty}
\label{annex:memstore}

The file \texttt{NOVAthesisFiles/StyFiles/memstore.sty} is a modern and robust memory storage engine for \emph{NOVAthesis}. Built using \LaTeX3 (\texttt{expl3}), it leverages property lists (\texttt{prop}) to store and retrieve data with advanced features like automatic space trimming and expansion control.

\section{Core Commands}
The primary commands are designed to be fast and reliable:
\begin{itemize}
    \item \textbf{\textbackslash MemSet}: Stores a value under a specific key in a named storage container. It automatically trims leading and trailing spaces from the value.
    \item \textbf{\textbackslash MemGet}: An expandable command to retrieve a stored value. It includes error reporting for missing keys or storages.
    \item \textbf{\textbackslash NewMemStore}: A dynamic interface generator that creates custom commands like \texttt{\textbackslash SetAuthors} and \texttt{\textbackslash GetAuthors} for specific storage types.
\end{itemize}

\section{Conditional Logic}
The package provides a comprehensive set of logic helpers for conditional formatting:
\begin{itemize}
    \item \textbf{\textbackslash IfMemDefinedTF}: Checks if a key exists and executes different code based on the result.
    \item \textbf{\textbackslash IfMemBlankTF}: Checks if a value is blank (empty or just spaces), which is particularly useful for detecting optional fields like ``co-supervisor.''
    \item \textbf{\textbackslash IfMemVoidTF}: A ``smart'' conditional that returns true if a key is either missing or its value is blank.
\end{itemize}

\section{Advanced Iteration}
\texttt{memstore.sty} includes powerful tools for bulk data processing:
\begin{itemize}
    \item \textbf{\textbackslash MemForEach}: Iterates over every key-value pair in a storage container, allowing the template to automatically generate lists of authors, committees, or school-specific metadata.
    \item \textbf{\textbackslash MemMatch}: A searching tool that finds the first existing key from a list and stores its contents for immediate use.
\end{itemize}

\section{Summary}
In summary, \texttt{memstore.sty} is the state-of-the-art data engine for \emph{NOVAthesis}. It provides the underlying infrastructure for managing complex document metadata with the safety and performance of \texttt{expl3}.
