%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% fix-bib.tex
%% NOVA thesis documentation file
%%
%% Detailed explanation of fix-bib.bib functionality.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE fix-bib.tex}

\chapter{Explanation of fix-bib.bib}
\label{annex:fix-bib}

The file \texttt{NOVAthesisFiles/StyFiles/fix-bib.bib} contains a small set of mandatory or highly recommended bibliography entries for use with the \emph{NOVAthesis} template. These entries are often referenced within the template's own manual and example chapters.

\section{Template Documentation}
The most important entry in this file is \texttt{novathesis-manual}.
\begin{itemize}
    \item \textbf{Purpose}: It provides the official reference for the \emph{NOVAthesis} template itself.
    \item \textbf{Requirement}: The template's \texttt{template.tex} file explicitly requests that users \texttt{\textbackslash cite\{novathesis-manual\}} somewhere in their document as a way to acknowledge the work put into the template.
\end{itemize}

\section{\LaTeX{} and AI Disclosure References}
Additional entries provide standard citations for foundational technologies and emerging academic requirements:
\begin{itemize}
    \item \textbf{latex-project}: A generic citation for the \LaTeX{} document preparation system.
    \item \textbf{Suchikova:2025:GAIDeT}: A reference to the ``Generative AI Delegation Taxonomy'' (GAIDeT). This is useful when documenting the use of AI in scientific research, as required by modern academic ethics.
    \item \textbf{Lourenco:2025:aidisclose}: A reference to the \texttt{aidisclose} package, which is used within \emph{NOVAthesis} to generate AI disclosure statements and checklists.
\end{itemize}

\section{Technical Note}
This \texttt{.bib} file is typically included in the project by the \texttt{\textbackslash ntaddfile\{bib\}\{fix-bib.bib\}} command in the configuration files. This ensures that these specific keys are always available for citation regardless of which other bibliography databases the user may be using.

\section{Summary}
The \texttt{fix-bib.bib} file ensures that critical references for the template, \LaTeX, and AI ethics are always available and correctly formatted for any document produced with \emph{NOVAthesis}.
