%!TEX root = ../template.tex
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% novathesis-spine.tex
%% NOVA thesis documentation file
%%
%% Detailed explanation of novathesis-spine.sty functionality.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\typeout{NT FILE novathesis-spine.tex}

\chapter{Explanation of novathesis-spine.sty}
\label{annex:novathesis-spine}

The file \texttt{NOVAthesisFiles/StyFiles/novathesis-spine.sty} provides a highly configurable and modern engine for generating thesis book spines. It is built using the \LaTeX3 (\texttt{expl3}) programming layer, \texttt{TikZ}, and \texttt{tcolorbox}.

\section{Core Commands}
Two main commands are provided for the user:
\begin{itemize}
    \item \textbf{\textbackslash NTSetup}: Used to configure spine parameters such as text, colors, sizes, and images via a key-value interface.
    \item \textbf{\textbackslash PrintBookSpine}: Typesets a single page containing the horizontal spine strip, automatically adjusted to the required width and thickness.
\end{itemize}

\section{Configurable Features}
The spine generator is extremely flexible, allowing for:
\begin{itemize}
    \item \textbf{Box-Based Layout}: The spine is divided into logical boxes (typically date, title, author, and logo). The order and length of these boxes can be freely rearranged via the \texttt{spine/order} key.
    \item \textbf{Dynamic Fitting}: It uses the \texttt{tcolorbox} ``fitting'' library to automatically scale text (like long thesis titles) to fit within their designated boxes.
    \item \textbf{Backgrounds}: Both the entire spine and individual boxes can have custom background colors or images (e.g., school logos or watermarks).
    \item \textbf{Sanitization}: The \texttt{\textbackslash NTSanitizeTitle} command ensures that spaces in titles are properly handled (using non-breaking spaces) to avoid awkward line breaks on the thin spine.
\end{itemize}

\section{Implementation Details}
The engine works by drawing a full-page \texttt{TikZ} overlay. It temporarily adjusts the paper (stock) size so that the page height matches the requested spine thickness, ensuring that the generated PDF can be sent directly to a printer for the book cover.

\section{Summary}
In summary, \texttt{novathesis-spine.sty} is a professional-grade tool that automates one of the most tedious parts of thesis preparation, ensuring that the final printed version looks polished from every angle.
